Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-22 WSSSPE1_report/WSSSPE1_summary_report.tex
Minor changes.

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center (i.e., a place to `take care' of software
when the original developers have stopped doing so).
The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

On the other hand, utility software presents opportunities to be usable by a
larger community making its undertaking and development an attractive pursuit.
It is generally more visible in community which in turn leads to a broader and
deeper participation. For instance, promoting collaborations across the breadth
(e.g.  different departments) and depth (e.g. stakeholders within a department)
of community, one of the key ingredient of a sustainable process. Successful
utility projects reap high rewards and have a longer usage span. Development
process becomes user-driven and self-sustaining.

@@

Finally, it appears that the career options of developers from different kinds
of software projects differ significantly. In order to further investigate the
much discussed issue of career paths for developers of these software, we
conducted a small survey of the cases discussed above with respect to the
current status of its core developers. We looked at the workshop submissions
and the original cited papers for the respective software. Where available, we
looked at the web pages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and excluded
relatively new projects such as \toolname{SciPy} and \toolname{numFOCUS} from
this survey. We find that in a majority of scientific software, more than one
original developer are still around, in some cases after more than 10 years,
(eg. \toolname{R/qtl}).  Whereas, in the case of utility software projects, it
appears that at least one original developer is still around, often in a
part-time or ``benefactor'' role. In other cases, it appears that the original
developers have went on to assume a larger role, with the same project.

>>>>

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center (i.e., a place to `take care' of software
when the original developers have stopped doing so).
The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g., Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

On the other hand, utility software presents opportunities to be usable by a
larger community making its undertaking and development an attractive pursuit.
It is generally more visible in community which in turn leads to a broader and
deeper participation. For instance, promoting collaborations across the breadth
(e.g., different departments) and depth (e.g., stakeholders within a department)
of community, one of the key ingredient of a sustainable process. Successful
utility projects reap high rewards and have a longer usage span. Development
process becomes user-driven and self-sustaining.

@@

Finally, it appears that the career options of developers from different kinds
of software projects differ significantly. In order to further investigate the
much discussed issue of career paths for developers of these software, we
conducted a small survey of the cases discussed above with respect to the
current status of its core developers. We looked at the workshop submissions
and the original cited papers for the respective software. Where available, we
looked at the web pages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and excluded
relatively new projects such as \toolname{SciPy} and \toolname{numFOCUS} from
this survey. We find that in a majority of scientific software, more than one
original developer are still around, in some cases after more than 10 years
(e.g., \toolname{R/qtl}).  Whereas, in the case of utility software projects, it
appears that at least one original developer is still around, often in a
part-time or ``benefactor'' role. In other cases, it appears that the original
developers have went on to assume a larger role, with the same project.
